Efficient pedigree recording for fast population genetics simulation

### 2017-11-20 forwards_paper.tex

\jda{Is this the right place for this paragraph? should it maybe be integrated
	with graf beginning 'Modern computing' above?}
\plr{this paragraph is supposed to cite the fwds time simulation methods we need to - @molpopgen fix this up?}
Forwards-time population genetics simulations are growing in importance,
as the genome-wide impacts of hitchhiking and background selection are more widely appreciated.
For instance, \citet{harris2016genetic} used SLiM \citep{slim} to simulate tens of thousands of human exomes
to assess the impact of genetic load and Neanderthal introgression on human genetic diversity.
This is roughly state-of-the-art
-- the fastest simulation frameworks such as SLiM and fwdpp \citep{fwdpp}
can ``only'' simulate tens of megabases of sequence in tens of thousands of individuals
for tens of thousands of generations.
This is still substantially short of whole genomes (hundreds to thousands of megabases)
for many biological population sizes of interest
(e.g., millions for a restricted population of mosquitos).
Achieving these sizes will likely require parallelization \citep{lawrie2017accelerating}.
The general idea of storing genealogical information to speed up simulations is not new --
it was entirely implemented in AnA-FiTS \citep{aberer2013rapid}, for instance --
but this program did not implement the critical step of discarding irrelevant genealogical information.
A similar but more limited method for discarding this information does appear in \citet{padhukasahasram2008exploring}.

>>>>

\jda{Is this the right place for this paragraph? should it maybe be integrated
	with graf beginning 'Modern computing' above?}
\plr{this paragraph is supposed to cite the fwds time simulation methods we need to - @molpopgen fix this up?}
Forwards-time population genetics simulations are growing in importance,
as the genome-wide impacts of hitchhiking and background selection are more widely appreciated.
For instance, \citet{harris2016genetic} used SLiM \citep{slim} to simulate ten thousand human exomes
to assess the impact of genetic load and Neanderthal introgression on human genetic diversity. 
\cite{Sanjak2017-ko} used
fwdpp \citep{fwdpp} to simulate a series of models of quantiative traits under mutation-selection balance with
population sizes of $2 times 10^4$ diploids in stable populations and populations growing up to $\approx 5
\times 10^5$ individuals, using the output to explore the relationship between the genotype/phenotype model and GWAS
outcomes.

Due to tracking neutral mutations, the fastest simulation frameworks such as SLiM, SLiM2 \cite{haller2017flexible} and fwdpp \citep{fwdpp}
can ``only'' simulate tens of megabases of sequence in tens of thousands of individuals
for tens of thousands of generations, and there are still near-order-of-magnitude performance differences between these
implementations \cite{haller2017flexible}. In practice, current state-of-the-art simulation software may take on the
order of 
weeks to simulate models without selection\cite{fwdpp,Hernandez2015-wf}, and different simulation engines differ in how efficiently they
calculate fitnesses in models with selection \cite{fwdpp}.
These population and region sizes are still substantially short of whole genomes (hundreds to thousands of megabases)
for many biological population sizes of interest.
Simulating very large popluations and entire genomes will likely require parallelization, which \cite{lawrie2017accelerating} used to improve
performance for the "Poisson Random Field" family of models \cite{Sawyer1992-jw}, in which mutations are unlinked and there is no
effect of genetic background on fitness.  However, little, and perhaps even no, attention has been paid to
parallelization of simulations involving linkage.  Here, we show that recording genealogical information instead of
explicitly tracking neutral mutations results in order-of-magnitude performance improvements to simulations based on
\fwdpp{}.  The  idea of storing genealogical information to speed up simulations is not new.  AnA-FiTS implemented it
completely \citep{aberer2013rapid}, but this program did not implement the critical step of discarding irrelevant genealogical information.  A similar but more limited method for discarding this information does appear in \citet{padhukasahasram2008exploring}.
